On the eletrodynamic of moving bodies

### 2019-08-09 Article_AS.Rmd
Mat et met

During the first half of the 20^th^ century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Hence, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties of commercial species such as white spruce. For example, improvement programs by genetic selection are in place for white spruce *(Picea glauca [Moench] Voss)* [@DRF2000; @Mullin2011] which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015] in North America. In Quebec, several generations of seed orchards have already been established in white spruce. Thus, the current objectives of genetic selection for this species are to select trees with a higher wood quality because white spruce has already reached a level considered interesting for traits such as rapid growth and volume of wood. \

@@

The material used was sampled from an experimental provenance-progeny test plantations designed to assess the genetic merit of first generation selections, which had been subdivided into breeding groups and delineated by provenance or geographic region. The details of the setting up of the experiment are described in @beaulieu1996. Crosses were made according to a partial diallel mating design within each of the six different breeding groups or sublines [@Burdon1983], which were designed to limit future inbreeding to within group. Each parent tree has been used in crossings one to three times, giving half-brother families (trees with one of two parents in common) and full-brother families (trees with both parents in common) in breeding groups. Genetic testing were set up in 1999 on 2-year-old seedlings from these experimental plantations near Asselin balsam fir-yellow birch ecological domain (cool weather), and Saint-Casimir located in the maple–basswood ecological domain (warm weather), Quebec, Canada, in randomized complete blocks to avoid any possible impact of heterogeneity within each site (table 1). Trees were assigned to plots in rows of five trees / plot (spacing of 2 m × 2 m) [@beaulieu2014]. In this study, we were used 93 full-brother families for a total of 2365 white spruce trees. The sampled trees were drawn from 18 blocks over the two sites and 9–14 individuals were sampled per full-sib cross. Trees used in this study are exclusively full-brother families.\

@@

Wood cores were extracted to a thickness of 1,68 mm from the south facing side of each tree 15-year-old trees in 2015 with a breast-height diameter (DBH15, 1.3 m above ground). Tree cores are extracted using increment borer which is a specialized tool for extracting wood cores from a living tree with relatively minor damage to the plant itself. These cores were used to determine ring width and wood density. Density was determined by scanning incremental cores from pith-to-bark along the radial face at a resolution of 25 $\mu$m in the QTRS-01X ring analyzer (Quintek Measurement Systems Inc., Knoxville, TN, USA).  A weighted mean density line was obtained by summing the product between the annual density measurements and the ratio of the respective radial areas to the total radial area. Wood cores were conditioned to about 7% moisture before scanning. Thus, the determination of the density of the wood makes it possible to calculate a quality index of it. From the density profiles, the limits of the annual growth rings were determined together with the relative proportions of earlywood and latewood in each annual ring using in-house data processing routines. Ring width determined through information of densitometry profiles. The transition between earlywood and the latewood is determined using the "Unix" script. This script used a transition density threshold ($D_t$) determined for each ring as follows:

@@

As part of this study, it is essential to accurately identify the year of formation of the annual rings measured. To this end, we used the technique of Cross-dating which consists in obtaining the concordance of the number of rings and of their width pattern to avoid counting errors or those caused by missing rings or by false annual rings. Crossdating is the most basic principle of dendrochronology. Cross-dating is a technique that ensures each individual tree ring is assigned its exact year of formation. This is accomplished by matching patterns of wide and narrow rings between cores from the same tree, and between trees from different locations.
To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983]. Thus, as the annual ring separation was carried out by an algorithm, we eliminated the samples (127 samples) whose dating was not correct and could not be corrected. Any incomplete or false rings were excluded from the dataset.

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce for the purpose of modeling the known density variation portion and subtracting it to better observe the rest of the density variations. Considering the hierarchical structure of the data, with observations on annual rings nested in the cores, cores nested in the trees, and trees nested in blocks, we opted for a mixed-effects modelling approach using nlme package [@NLME2018] in the R statistical programming environment [@R2018] (Eq.3). Previous studies have used this method to describe the density profile of different spruce species that have a similar profile [@franceschini2010; @gardiner2011]. 
However, when modeling density profiles we found that the variation between individual discs follows different patterns from the pith-to-bark between tree. We noticed that it was impossible to optimize the equation using the same estimated parameters for all tree cores, because the pattern of variation of annual rings average density of each tree is more or less different from the others. To work around this problem, we chose to model each of the profiles separately. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core . The fit of the model was evaluated by visual analysis of the model residual curves against adjusted values in addition to the explanatory variables, fit indices (R^2^) [@Pape1999].
As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

where RD is the mean ring density (kg.m^-3^), CA is cambial age (years), $b_0$, $b_1$ and $b_2$ are empirically determined parameters, and $e$ is the residual error term. Parameter $b_0$ denotes the intercept when CA approaches a theoretical value of 0. The graph in figure 2 describes the three parameters of our model and allows us to understand how each parameter influences our model (Fig 2).

@@

## Determination of interannual variation amplitude of wood density index
After subtraction of the typical density, it is observed that the residues remain variable. Knowing that the low frequency signal is subtracted, this variation can only be due to the high frequency signal which is the interannual variation. As explained earlier in the introduction, low interannual variation is an important criterion for wood quality. In order to identify tree families with low interannual variation amplitude in density, we have created a viability index.
The interannual variation amplitude of wood density index is an index to determine the variability in the density of pith-to-bark in order to identify tree families that are least sensitive to variations in climatic conditions. The residues of our model were used to obtain the interannual variation amplitude of wood density index, which was determined using an equation. We subtracted the density residues values of year $n-1$ from year $n$. We were interested more specifically in the difference between the average density residues of annual rings of two consecutive years, no matter whether the difference was positive or negative. The absolute value was used for each difference to get rid of the sign. The index was estimated from the following formula:

@@

Where V is the interannual variation amplitude of wood density index, $D_n$ is the wood density residues of year $n$ and  $D_{n - 1}$ is the wood density residues of year $n-1$.

@@

## Quantitative genetic analyses
Variance and covariance for genetics were estimated using mixed models [@Lenz2013] of the R software. The equation used is as follows:

$$ y_{ijkl} = \mu + s_i + b_{(i)j} + f_k + bf_{(i)jk} + e_{ijkl}$$

where $y_{ijkl}$  is the observation on the $ijklth$ tree, $\mu$ is the overall mean, $s_i$ is the fixed site effect, $b_{(i)j}$ is the fixed effect of the $j^{th}$ bloc in site $i$, $f_k$ is the random effect of the $k^{th}$ family, $bf_{(i)j}$ is the random interaction of family and bloc, and $e_{ijkl}$ is the residual error.
In a full-sib design with random mating, members of the same family are presumed to share half of their genes [@white2007]. Therefore, individual heritability was obtained as follows:

@@

The model we have developed has made it possible to optimize the density of the wood for each core. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. For the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\

@@

![Pith-to-bark radial pattern of breast height density by ring. Fine lines show densities for individual stands of the samples for the Asselin site.The bold lines shows the density predicted by Model](graph_table/Density_asselin.png){width=330px; height=330px} \

![Pith-to-bark radial pattern of breast height density by ring. Fine lines show densities for individual stands of the samples for the Saint-Casimir site.The bold lines shows the density predicted by Model](graph_table/Density_casimir.png){width=330px; height=330px} \

When considering all the cores of trees from the pith-to-bark, ring density was high near the pith and decreased rapidly in the first 5 growth rings. It goes from 430 kg/m^3^ to 370 kg/m^3^ for the Saint-Casimir site and from 455kg/m^3^ to kg/m^3^ for the Asselin site. Then, this decrease was followed by a slight increase until the twelfth year. \

Model predictions are similar to data trends across sites (fig 3&4). A decrease in density is observed in the first years of growth and then an increase in density from a certain cambial age. Whether it is the data or the predictions of the model, we observe a difference between the Asselin (fig 3) and Saint-Casimir (fig 4) sites. The increase in density after the decrease of the first years of growth is greater in Saint-Casimir than in Asselin. However, the average density is very similar between the two sites. In spite of the interannual variations, the trees have the same general tendency and this one seems well described by the model which adapts to each of the sites.\

@@

|Site                     |Min  |Mean  |Max    |
|:------------------------|----:|-----:|------:|
|Asselin (Témiscouata)    |  0.0| 417.5| 1155.8|
|Saint-Casimir (Portneuf) | 54.9| 339.8|  746.0|
|Combined                 |  0.0| 375.9| 1155.8|
Table: Summary table of variability index

When comparing the index of variability between study sites, there is greater variability in Asselin than in Saint-Casimir. The maximum value of 1155.8 in Asselin, for example, is 746, which represents a huge difference.

@@

The results of the rank correlation coefficient are summarized in Table 3. These results show an absence of correlation for the variability index (p-value = 0.43) and a strong correlation for all the parameters of the model (p-value = 2.2 10^-16^). The value of Rho is very similar between the three parameters.

>>>>

During the first half of the 20^th^ century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Hence, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties of commercial species such as white spruce. For example, improvement programs by genetic selection are in place for white spruce (*Picea glauca* [Moench] Voss) [@DRF2000; @Mullin2011] which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015] in North America. In Quebec, several generations of seed orchards have already been established in white spruce. Thus, the current objectives of genetic selection for this species are to select trees with a higher wood quality because white spruce has already reached a level considered interesting for traits such as rapid growth and volume of wood. \

@@

The material used was sampled from an experimental provenance-progeny test plantations designed to assess the genetic merit of first generation selections, which had been subdivided into breeding groups and delineated by provenance or geographic region. The details of the setting up of the experiment are described in @beaulieu1996. Crosses were made according to a partial diallel mating design within each of the six different breeding groups or sublines [@Burdon1983], which were designed to limit future inbreeding to within group. Each parent tree has been used in crossings one to three times, giving half-brother families (trees with one of two parents in common) and full-brother families (trees with both parents in common) in breeding groups. Genetic testing were set up in 1999 on 2-year-old seedlings from these experimental plantations near Asselin balsam fir-yellow birch ecological domain (cool weather), and Saint-Casimir located in the maple–basswood ecological domain (warm weather), Quebec, Canada, in randomized complete blocks to avoid any possible impact of heterogeneity within each site (table 1). Trees were assigned to plots in rows of five trees / plot (spacing of 2 m × 2 m) [@beaulieu2014]. In this study, we were used 93 full-brother families for a total of 2365 white spruce trees. The sampled trees were drawn from 18 plots over the two sites and 9–14 individuals were sampled per full-sib cross. Trees used in this study are exclusively full-brother families.\

@@

Wood cores were extracted to a thickness of 1,68 mm from the south facing side of each tree 15-year-old trees in 2015 with a breast-height diameter (DBH15, 1.3 m above ground). Tree cores are extracted using increment borer which is a specialized tool for extracting wood cores from a living tree with relatively minor damage to the plant itself. These cores were used to determine ring width and wood density. Density was determined by scanning incremental cores from pith-to-bark along the radial face at a resolution of 25 $\mu$m in the QTRS-01X ring analyzer (Quintek Measurement Systems Inc., Knoxville, TN, USA). A weighted mean density line was obtained by summing the product between the annual density measurements and the ratio of the respective radial areas to the total radial area. Wood cores were conditioned to about 7% moisture before scanning. Thus, the determination of the wood density makes it possible to calculate a quality index of it. From the density profiles, the limits of the annual growth rings were determined together with the relative proportions of earlywood and latewood in each annual ring using in-house data processing routines. Ring width determined through information of densitometry profiles. The transition between earlywood and the latewood is determined using the "Unix" script. This script used a transition density threshold ($D_t$) determined for each ring as follows:

@@

As part of this study, it is essential to accurately identify the year of formation of the annual rings measured. To this end, we used the technique of Crossdating which consists in obtaining the concordance of the number of rings and of their width pattern to avoid counting errors or those caused by missing rings or by false annual rings. Crossdating  is one of the most important basic principles of dendrochronology. This technique ensures that each tree ring is assigned to the exact year of its formation. This is accomplished by matching patterns of wide and narrow rings between cores from the same tree, and between trees from different positions.
To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for crossdating and quality control of annual ring measurements [@HOLMES1983]. Thus, as the annual ring separation was carried out by an algorithm, we eliminated the samples (127 samples) whose dating was not correct and could not be corrected. Any incomplete or false rings were excluded from the dataset.

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce for the purpose of modeling the known density variation portion and subtracting it to better observe the rest of the density variations. Considering the hierarchical structure of the data, with observations on annual rings nested in the cores, cores nested in the trees, and trees nested in blocks, we opted for a mixed-effects modelling approach using nlme package [@NLME2018] in the R statistical programming environment [@R2018]. Previous studies have used this method to describe the density profile of different spruce species that have a similar profile [@franceschini2010; @gardiner2011]. 
However, when modeling density profiles we found that the variation between individual cores follows different patterns from the pith-to-bark between trees. We noticed that it was impossible to optimize the equation using the same estimated parameters for all tree cores, because the pattern of variation of annual rings average density of each tree is more or less different from the others. To circumvent this problem, we chose to model each of the profiles separately. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core. Model fit of the model was evaluated by visual analysis of the model residual curves against adjusted values in addition to the explanatory variables, fit indices (R^2^) [@Pape1999].
As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

where RD is the mean ring density (kg.m^-3^), CA is cambial age (years), $b_0$, $b_1$ and $b_2$ are empirically determined parameters, and $e$ is the residual error term. Parameter $b_0$ denotes the intercept when CA approaches a theoretical value of 0; $b_1$ denotes the scale parameter of the early decrease in RD. $b_2$ denotes the rise in RD at the end of the juvenile wood and at the beginning of the mature wood formation. The error term $e$ was assumed to be normally distributed, with N (0, $\sigma^{2}$). The first exponential term $b_0 \times \exp\left(\frac{b_1}{CA}\right)$ has a value which decreases rapidly in the first few annual rings and then gradually approaches an asymptote close to zero. The graph in figure 2 describes the three parameters of our model and allows us to understand how each parameter influences our model (Fig 2).

@@

## Determination of interannual radial variation amplitude of wood density index
After subtraction of the typical density, it is observed that the residues remain variable. Knowing that the low frequency signal is subtracted, this radial variation can only be due to the high frequency signal which is the interannual radial variation. As explained earlier in the introduction, low interannual radial variation is an important criterion for wood quality. In order to identify tree families with low interannual radial variation amplitude in density, we have created a interannual radial variation amplitude index.
The interannual variation amplitude of wood density index is an index to determine the variability in the density of pith-to-bark in order to identify tree families that are least sensitive to variations in climatic conditions. The residues of our model were used to obtain the interannual variation amplitude of wood density index, which was determined using an equation. We subtracted the density residues values of year $n-1$ from year $n$. We were interested more specifically in the difference between the average density residues of annual rings of two consecutive years, no matter whether the difference was positive or negative. The absolute value was used for each difference to get rid of the sign. The index was estimated from the following formula:

@@

Where V is the interannual radial variation amplitude of wood density index, $D_n$ is the wood density residues of year $n$ and  $D_{n - 1}$ is the wood density residues of year $n-1$.

@@

## Quantitative genetic analyses
Variance and covariance for genetics were estimated using linear mixed-effects model [@Lenz2013] of the R software. The equation used is as follows:

$$ Y_{ijkl} = \mu + s_i + b_{(i)j} + f_k + bf_{(i)jk} + e_{ijkl}$$

where $Y_{ijkl}$  is the observation on the $ijklth$ tree, $\mu$ is the overall mean, $s_i$ is the fixed site effect, $b_{(i)j}$ is the fixed effect of the $j^{th}$ bloc in site $i$, $f_k$ is the random effect of the $k^{th}$ family, $bf_{(i)j}$ is the random interaction of family and bloc, and $e_{ijkl}$ is the residual error.
In a full-sib design with random mating, members of the same family are presumed to share half of their genes [@white2007]. Therefore, individual heritability was obtained as follows:

@@

The model we have developed has made it possible to optimize the density of the wood for each core. A summary of the parameter values obtained of the model fitting process (Eq. 1) can be found in Table 2. We observe that the average value of the model parameters differs between Asselin and Saint-Casimir. For the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin. Although there are small differences in the parameters between the two sites, these differences are very small. The fit of the model is therefore similar at both sites.\

@@

![Pith-to-bark radial pattern of breast height density by ring. Fine lines show densities for individual stands of the samples for the Asselin site.The bold lines shows the density predicted by Model](graph_table/Density_asselin.png){width=330px; height=330px}

![Pith-to-bark radial pattern of breast height density by ring. Fine lines show densities for individual stands of the samples for the Saint-Casimir site.The bold lines shows the density predicted by Model](graph_table/Density_casimir.png){width=330px; height=330px}

When considering all the cores of trees from the pith-to-bark, ring density was high near the pith and decreased rapidly in the first 5 growth rings. It declined from 430 kg/m^3^ to 370 kg/m^3^ for the Saint-Casimir site and from 455kg/m^3^ to 375 kg/m^3^ for the Asselin site. Then, this decrease was followed by a slow increase until the twelfth year. \

Model predictions are similar to data trends across sites (figs 4 and 5). A decrease in density is observed in the first years of growth and then an increase in density from a certain cambial age. Whether it is the data or the predictions of the model, we observe a difference between the Asselin (fig 4) and Saint-Casimir (fig 5) sites. The increase in density after the decrease of the first years of growth is greater in Saint-Casimir than in Asselin. However, the average density is very similar between the two sites. In spite of the interannual variations, the trees have the same general tendency and this one seems well described by the model which adapts to each of the sites.\

@@

|Site                     |Min  |Mean  |Max    |
|:------------------------|----:|-----:|------:|
|Asselin (Témiscouata)    |116.5| 417.9| 1155.8|
|Saint-Casimir (Portneuf) | 54.9| 339.8|  746.0|
|Combined                 | 54.9| 375.9| 1155.8|
Table: Summary table of variability index

When comparing the index of variability between study sites, there is greater variability in Asselin than in Saint-Casimir. A summary of the density interannual amplitude index values can be found in table 3. The maximum value is 1155.8 in Asselin, and 746 for the Saint-Casimir site, which represents a huge difference. \

@@

The results of the rank correlation coefficient are summarized in Table 4. These results show an absence of correlation for the variability index (p-value = 0.43) and a strong correlation for all the parameters of the model (p-value = 2.2 10^-16^). The value of Rho is very similar between the three parameters. \
